Detecting Hate Speech with GPT-3

### 2022-02-25 outputs/documents/paper.Rmd
Updates to results section

@@

We use all of the 998 statements in the ETHOS dataset that have a binary classification of hate speech or not hate speech. Of these, the 433 statements that contain hate speech additionally have multiple labels that classify the content, for instance, does the comment have to do with violence, gender, race, nationality, disability, etc. We considered all of the 136 statements that contain race-based hate speech, and we focus on the 76 whose race-based score is at least 0.5. Similarly, we considered the all of 174 statements that contain gender-based hate speech, and again focused on the 84 whose gender-based score is at least 0.5. To create a balanced dataset we randomly draw 120 of these statements, using stratified sampling, to result in 60 sexist statements and 60 racist statements. For comparison purposes, we randomly draw 120 statements that are classified as not hate speech. The balanced sample enables us to more easily understand our results.

@@

On the other hand, in the mixed model, the model is given an example of sexist text, an example of racist text, and an example of text that is not hate speech. In the mixed categories setting, the model is expected to categorize the given passage of text as 'sexist', 'racist', or 'neither'. To simplify the classification procedure and allow for comparisons between different example inputs, we created ten sets of examples against which 240 comments were classified each. We elected to use ten example sets to allow for some variation in results while keeping the number of groupings and size of the resulting data set reasonably small.  The example sets each consist of one sexist, one racist, and one 'neither' comment and were created using the comments at indexes one through ten from each dataset. Each set of examples is used for the model to classify 60 racist comments, 60 sexist comments, and 120 comments that are in 'neither' category. The inputted comments were the first _n_ + 1 comments from their respective data sets with the _i_^th comment dropped corresponding to the use of example set _i_. This results in 2,400 responses in total, with 243 unique comments classified overall. The ten sets of examples are available in Table \@ref(tab:appendixa) in Appendix A. An example of this is:

@@

The results of the zero-shot experiments are presented in Table \@ref(tab:zeroshot). The model has 35 matches and 25 mismatches in the sexist category, and 23 matches and 37 mismatches in the racist category. The model performs better when identifying sexist comments compared with identifying racist comments. However, the overall ratio of matches and mismatches is 58:62. In other words, the accuracy in identifying hate speech in the zero-shot setting is 48 per cent.

@@

```{r zeroshot-racism}
zero_shot_result <- read_csv(here::here("outputs/data/zero_shot_results.csv"), show_col_types = FALSE)

@@

The results of the one-shot learning experiments are presented in Table \@ref(tab:oneshot). The model has 46 matches and 14 mismatches in the racist category, and 37 matches and 23 mismatches in the sexist category. In contrast with the result generated from zero-shot learning, the model performs slightly better in identifying racist comments compared with identifying sexist comments. The general performance in the one-shot setting is also better than in the zero-shot setting. The overall ratio of matches and mismatches is 83:37. In other words, the accuracy of identifying hate speech in the one-shot setting is 69.2 per cent.

@@

The results of the single-category, few-shot learning, experiments are presented in Table \@ref(tab:fewshotsingle). The model has 41 matches and 19 mismatches in the racist category, and 42 matches and 18 mismatches in the sexist category. Here, the model performs almost equally well at identifying racist comments compared with identifying sexist comments. The general performance in the few-shot learning setting is similar to the performance in the one-shot learning setting. The overall ratio of matches and mismatches is 83:37, or around 69.2 per cent, however the composition is different, compared with the one-shot learning setting.

@@

The results of the mixed-category few-shot experiments are presented in Table \@ref(tab:fewshotmixed). Among the ten sets of examples, Example Set 2 yields the best performance, as the model has the highest number of matches, 180, and the lowest number of mismatches, 60. In other words, the accuracy under the mixed category few-shot setting with Example Set 2 is 75 per cent. The example set that yields the worst results is Example Set 9, in which there are 137 matches and 103 mismatches. The accuracy rate with Example Set 9 is 57.1 per cent. The difference between Example Sets 2 and 9 suggests that, although the models are provided with same number of examples, the content of the examples also affects how the model makes inferences.

@@

@@

To reduce the chance of the model generating answers that are out of scope, a brief instruction is added to the prompt, specifying that the answers be: 'sexist', 'racist', or 'neither'. The addition of an instruction successfully restricts the generated answers within the specified terms. The unique generated answers are: 'racist', 'neither', and 'sexist'.

@@

The results of the mixed-category, few-shot learning, with instruction, experiments are presented in Table \@ref(tab:fewshotmixedinstruct). With the addition of an instruction in the prompt, the example set that yields the best result is Example Set 1 instead of Example Set 2. The addition of the instruction in the prompt increases the highest number of matches from 180 to 187; the highest accuracy rate is increased from 75 per cent to 78 per cent. In almost all cases the accuracy of the model increases (Figure \ref{fig:comparison}). Across all the example sets, there are 1,785, or 74 per cent, matches and 615, or 26 per cent, mismatches.

@@

# Example sets
few_shot_mixed_result_instruct_racism <- 
  few_shot_instruction_result %>%
  mutate(example_set = example_set + 1,
         label = case_when(!grepl("racist", label) ~ "not racist", TRUE ~ "racist"),
         answer = case_when(!grepl("racist", answer) ~ "not racist", TRUE ~ "racist")) %>%
  count(example_set, label, answer) %>%
  pivot_wider(names_from = answer, values_from = n) %>%
  mutate(label = case_when(label == "not racist" ~ "Not racist",
                           label == "racist" ~ "Racist")) %>%
  rename("Actual classification" = label, "Not racist" = "not racist", "Racist" = "racist")

@@

few_shot_instruction_result <- 
  few_shot_instruction_result %>% 
  rename(Result = result,
         "Example set" = example_set,
         Correct = result_cnt) %>% 
  mutate(Type = "With instruction")

>>>>

<!-- Insert brief description of GPT-3 - the general usage of the model, its transformer-based architecture, what are the inputs-outputs, etc. -->

@@

We use all of the 998 statements in the ETHOS dataset that have a binary classification of hate speech or not hate speech. Of these, the 433 statements that contain hate speech additionally have multiple labels that classify the content, for instance, does the comment have to do with violence, gender, race, nationality, disability, etc. We considered all of the 136 statements that contain race-based hate speech, and we focus on the 76 whose race-based score is at least 0.5. Similarly, we considered the all of 174 statements that contain gender-based hate speech, and again focused on the 84 whose gender-based score is at least 0.5. To create a balanced dataset we select 120 of these statements to result in 60 sexist statements and 60 racist statements. For comparison purposes, we randomly draw 120 statements that are classified as not hate speech. The balanced sample enables us to more easily understand our results.

@@

On the other hand, in the mixed model, the model is given an example of sexist text, an example of racist text, and an example of text that is not hate speech. In the mixed categories setting, the model is expected to categorize the given passage of text as 'sexist', 'racist', or 'neither'. To simplify the classification procedure and allow for comparisons between different example inputs, we created ten sets of examples against which 240 comments were classified each. We elected to use ten example sets to allow for some variation in results while keeping the number of groupings and size of the resulting data set reasonably small.  The example sets each consist of one sexist, one racist, and one 'neither' comment and were created using the comments at indexes one through ten from each dataset. Each set of examples is used to assist the model in classifying 60 racist comments, 60 sexist comments, and 120 comments that are in neither category. The inputted comments were the first _n_ + 1 comments from their respective data sets with the _i_^th^ comment dropped corresponding to the use of example set _i_. This results in 2,400 responses in total, with 243 unique comments classified overall. The ten sets of examples are available in Table \@ref(tab:appendixa) in Appendix A. An example of this is:

@@

The results of the zero-shot experiments are presented in Table \@ref(tab:zeroshot-racism), Table \@ref(tab:zeroshot-sexism), Table \@ref(tab:zeroshot-hate), and \@ref(tab:zeroshot-summary). We assess GPT-3's performance in all settings using accuracy, precision, recall, and F1 score. Accuracy is the proportion of correctly classified comments (hate speech and non-hate speech) out of all comments classified. Precision is the proportion of hate speech comments correctly identified out of all comments identified as hate speech (both correctly and incorrectly). Recall is the number of hate speech comments correctly identified out of all hate speech comments in the data set (both correctly and incorrectly identified). F1 score is the harmonic mean of precision and recall. In the case of hate speech identification it is desirable to have a model with high recall, meaning a model that can identify a relatively high proportion of the hate speech text within a data set, however the F1 score can provide a more well-rounded metric for model performance and comparison.

The model has 34 matches (true positives and negatives) and 26 mismatches (false positives and negatives) in the sexist category, and 41 matches and 19 mismatches in the racist category. The model performs with higher accuracy and precision when identifying racist comments compared with identifying sexist comments. The overall ratio of matches and mismatches is 75:45. In other words, the accuracy in identifying hate speech in the zero-shot setting is 62.5 per cent.

@@

```{r zeroshot-racism, message=FALSE}
zero_shot_result <- read_csv(here::here("outputs/data/zero_shot_results.csv"), show_col_types = FALSE)

@@

The results of the one-shot learning experiments are presented in Table \@ref(tab:oneshot-racism), Table \@ref(tab:oneshot-sexism), Table \@ref(tab:oneshot-hate), and Table \@ref(tab:oneshot-summary). The model has 37 matches and 23 mismatches in the racist category, and 36 matches and 24 mismatches in the sexist category. In contrast with the result generated from zero-shot learning, the model approximately the same in identifying racist comments compared with identifying sexist comments. The general performance in the one-shot setting is worse than in the zero-shot setting, with lower accuracy, recall, and precision overall. The overall ratio of matches and mismatches is 73:47. In other words, the accuracy of identifying hate speech in the one-shot setting is 60.8 per cent.

@@

The results of the single-category, few-shot learning, experiments are presented in Table \@ref(tab:fewshotsingle-racism), Table \@ref(tab:fewshotsingle-sexism), Table \@ref(tab:fewshotsingle-hate), and Table \@ref(tab:fewshotsingle-summary). The model has 41 matches and 19 mismatches in the racist category, and 42 matches and 18 mismatches in the sexist category. Similar to the one-shot setting, the model performs almost equally well at identifying racist comments compared with identifying sexist comments. The general performance in the few-shot learning setting is similar to the performance in the one-shot learning setting as well. The overall ratio of matches and mismatches is 83:37, or around 69.2 per cent, with higher precision and lower recall compared with the one-shot learning setting.

@@

The results of the mixed-category few-shot experiments are presented in Table \@ref(tab:fewshotmixed-racism), Table \@ref(tab:fewshotmixed-sexism), and Table \@ref(tab:fewshotmixed-summary).

Among the ten sets of examples, Example Set 10 yields the best performance in terms of F1 score for both racist and sexist comments. It also has the highest accuracy for identifying sexist comments at 91.25 percent and is tied with Example Set 2 for the highest accuracy in identifying racist comments at 90.42 percent. The example set that yields the worst results in identifying racist text in terms of F1 score is Example Set 8, which has an F1 score of 68.26 percent (and the lowest accuracy at 77.92 percent) for this dataset. The example set that yields the worst results in identifying sexist text in terms of F1 score is Example Set 9, which has an F1 score of 68.45 percent (and the lowest accuracy at 80.42 percent) for this dataset. The differences between Example Sets 8, 9, and 10 suggest that, although the models are provided with same number of examples, the content of the examples also affects how the model makes inferences. Overall, the mixed-category few-shot setting performs slightly better at identifying sexist text versus racist text. It is also performs with higher accuracy and a higher F1 score than the zero-shot, one-shot, and single-category few-shot settings for both racist and sexist text.

@@

For the purposes of our analysis, a classification was considered a true positive if the answer outputted by GPT-3 contained a category that matched the comment's label. For example, if a comment was labelled 'sexist' and the comment was classified by the model as 'sexist, racist', this was considered a true positive. If a comment was labelled 'sexist' and the comment was classified by the model as 'racist', 'transphobic', 'neither', etc. this was considered a false negative.

@@

To reduce the chance of the model generating answers that are out of scope, a brief instruction is added to the prompt, specifying that the answers be: 'sexist', 'racist', or 'neither'. The addition of an instruction successfully restricts the generated answers within the specified terms. The unique generated answers are: 'racist', 'sexist', and 'neither'.

@@

The results of the mixed-category, few-shot learning, with instruction, experiments are presented in Table \@ref(tab:fewshotmixedinstruct-racism), Table \@ref(tab:fewshotmixedinstruct-sexism), and Table \@ref(tab:fewshotmixedinstruct-summary). With the addition of an instruction in the prompt, Example Set 10 remains the best performing example set in terms of accuracy (92.08 percent) and F1 score (84.03 percent) for sexist text, however Example Set 4 now performs best for racist text in terms of both accuracy (91.25 percent) and F1 score (83.20 percent).

<!-- Fix  below -->
The addition of the instruction in the prompt increases the highest number of matches from 180 to 187; the highest accuracy rate is increased from 75 per cent to 78 per cent. In almost all cases the accuracy of the model increases (Figure \ref{fig:comparison}). Across all the example sets, there are 1,785, or 74 per cent, matches and 615, or 26 per cent, mismatches.

@@

# Example sets
few_shot_mixed_result_instruct_racism <- 
  few_shot_instruction_result %>%
  mutate(example_set = example_set + 1,
         label = case_when(!grepl("racist", label) ~ "not racist", TRUE ~ "racist"),
         answer = case_when(!grepl("racist", answer) ~ "not racist", TRUE ~ "racist")) %>%
  count(example_set, label, answer) %>%
  pivot_wider(names_from = answer, values_from = n, values_fill = 0) %>%
  mutate(label = case_when(label == "not racist" ~ "Not racist",
                           label == "racist" ~ "Racist")) %>%
  rename("Actual classification" = label, "Not racist" = "not racist", "Racist" = "racist")

@@

few_shot_instruction_result <- 
  few_shot_instruction_result %>% 
  rename(Result = result,
         "Example Set" = example_set,
         Correct = result_cnt) %>% 
  mutate(Type = "With instruction")
